Finch: Sparse and Structured Array Programming with Control Flow

### 2024-04-05 sections/eval/spmv.tex
Update on Overleaf.

\begin{wrapfigure}{r}{0.41\linewidth}
    \vspace{-28pt}
    \begin{minipage}[t]{0.18\textwidth}
        \begin{minted}{julia}
            y .= 0
            for j = _, i = _
              y[i] += A[i, j] * x[j]
            end
        \end{minted}
        \vspace{24pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _, i = _
              y[j] += A[i, j] * x[i]
            end
        \end{minted}
    \end{minipage}\hfill%
    \begin{minipage}[t]{0.22\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _
              let x_j = x[j]
                y_j .= 0
                for i = _
                  let A_ij = A[i, j]
                    y[i] += x_j * A_ij
                    y_j[] += A_ij * x[i]
                  end
                end
                #D is the diagonal
                y[j] += y_j[] + D[j] * x_j
              end
            end
        \end{minted}
    \end{minipage}
    \vspace{-8pt}
    \caption{Finch row-major, column-major and symmetric SpMV Programs}
    \label{spmv_programs}
    \vspace{-12pt}
  \end{wrapfigure}
  \subsection{Sparse Matrix-Vector Multiply (SpMV)}
  Sparse matrix-vector multiplication (SpMV) has a wide range of applications and has been thoroughly studied~\cite{liu_csr5_2015,
  zhou_enabling_2020}. 
  %
  Because SpMV is bandwidth bound, many formats have
  been proposed to reduce the footprint~\cite{langr_evaluation_2016}. 
  %
  The wide range of applications unsurprisingly
  results in a wide range of array structures, making it an effective kernel to
  demonstrate the utility of our programming model. 
  %
  We varied both the data formats and the SpMV algorithm. 
  %
  Our formats
  are shown in Table~\ref{spmv_tensor_formats}, and our programs are listed in
  Figure~\ref{spmv_programs}.
  %
  In varying the program, we consider both row and column-major SpMV programs, as
  well as a symmetric SpMV. 
  %
  Finch enables us to exploit symmetry
  effectively.
  %
  Our program restricts our attention to the canonical triangle (using
  masks), reuses the reads to the canonical triangle of the symmetric matrix
  (using a $\finchdefine$ statement), and writes the results to both relevant
  locations (using multiple outputs).
  %
  All of our programs apply to all of our level formats, enabling exploitation of multiple structural patterns concurrently (e.g. sparsity \textit{and} symmetry).

  \begin{wrapfigure}{r}{0.5\linewidth}
    \scriptsize
    \centering
    \vspace{-12pt}
    \vspace{-12pt}
    \begin{tabular}{|r|l|l|}
        \hline
        \textbf{Format} & \textbf{Style of Matrix}\\
        \hline
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseList(Element(0.0)))} & real, sparse \\
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseList(Pattern()))} & boolean, sparse \\
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseVBL(Element(0.0)))} & real, blocked \\
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseBand(Element(0.0)))} & real, banded \\
        %\mintinline{julia}{Dense(SparsePoint(Element(0.0)))} & real-valued matrices with one value per row \\
        %\mintinline{julia}{Dense(SparseRLE(Pattern(0.0)))} & matrix with runs of true or false \\
        \hline 
    \end{tabular}
    \caption{SpMV Tensor Formats}
    \label{spmv_tensor_formats}
    \vspace{-8pt}
  \end{wrapfigure}

>>>>

\begin{wrapfigure}{r}{0.41\linewidth}
    \vspace{-36pt}
    \begin{minipage}[t]{0.18\textwidth}
        \begin{minted}{julia}

            y .= 0
            for j = _, i = _
              y[i] += A[i, j] * x[j]
            end
        \end{minted}
        \vspace{24pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _, i = _
              y[j] += A[i, j] * x[i]
            end
        \end{minted}
    \end{minipage}\hfill%
    \begin{minipage}[t]{0.22\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            y .= 0
            for j = _
              let x_j = x[j]
                y_j .= 0
                for i = _
                  let A_ij = A[i, j]
                    y[i] += x_j * A_ij
                    y_j[] += A_ij * x[i]
                  end
                end
                #D is the diagonal
                y[j] += y_j[] + D[j] * x_j
              end
            end
        \end{minted}
    \end{minipage}
    \vspace{-12pt}
    \caption{Finch row-major, column-major and symmetric SpMV Programs}
    \label{spmv_programs}
    \vspace{-36pt}
  \end{wrapfigure}
  \subsection{Sparse Matrix-Vector Multiply (SpMV)}
  Sparse matrix-vector multiply (SpMV) has a wide range of applications and has been thoroughly studied~\cite{liu_csr5_2015,
  zhou_enabling_2020}. 
  %
  Because SpMV is bandwidth bound, many formats have
  been proposed to reduce the footprint~\cite{langr_evaluation_2016}. 
  %
  The wide range of applications unsurprisingly
  results in a wide range of array structures, making it an effective kernel to
  demonstrate the utility of our programming model. 
  %
  We varied both the data formats and the SpMV algorithm. 
  %
  Our formats
  are shown in Table~\ref{spmv_tensor_formats}, and our programs are listed in
  Figure~\ref{spmv_programs}.
  %
  In varying the program, we consider both row and column-major SpMV programs, as
  well as a symmetric SpMV. 
  %
  Finch enables us to exploit symmetry
  effectively.
  %
  Our program restricts our attention to the canonical triangle (using
  masks), reuses the reads to the canonical triangle of the symmetric matrix
  (using a $\finchdefine$ statement), and writes the results to both relevant
  locations (using multiple outputs).
  %
  All of our programs apply to all of our level formats, enabling exploitation of multiple structural patterns concurrently (e.g. sparsity \textit{and} symmetry).

  \begin{wrapfigure}{r}{0.5\linewidth}
    \scriptsize
    \centering
    \vspace{-16pt}
    \begin{tabular}{|r|l|l|}
        \hline
        \textbf{Format} & \textbf{Style of Matrix}\\
        \hline
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseList(Element(0.0)))} & real, sparse \\
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseList(Pattern()))} & boolean, sparse \\
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseVBL(Element(0.0)))} & real, blocked \\
        \mintinline[fontsize=\scriptsize]{julia}{Dense(SparseBand(Element(0.0)))} & real, banded \\
        %\mintinline{julia}{Dense(SparsePoint(Element(0.0)))} & real-valued matrices with one value per row \\
        %\mintinline{julia}{Dense(SparseRLE(Pattern(0.0)))} & matrix with runs of true or false \\
        \hline 
    \end{tabular}
      \vspace{-8pt}
    \caption{SpMV Tensor Formats}
    \label{spmv_tensor_formats}
    \vspace{-8pt}
  \end{wrapfigure}
